---
title: "Untitled"
author: "RQ"
date: "14/03/2021"
output: pdf_document
bibliography: COVID-air-traffic.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r, libs, message=FALSE}
library(tidyverse)
library(sf)
library(rnaturalearth)

# own functions
my_funs <- list.files(path = "./R/", pattern = "\\.R", full.names = TRUE)
purrr::walk(my_funs, .f = source)

# local data storage
pth <- "../__DATA/OSN-COVID-ECTRL/"
```

## Overview

This Rmarkdown report documents the data preparatory steps for cleaning the Opensky Network data.

TODO

* describe data and data issues

## Data Ingestion

Opensky Network publishes on a weekly basis records of their global data (@osn_covid_dataset, @schafer_osn_2014, @olive_traffic_2019) .
These data is transferred as zipped csv files.
Given the file size the data is read in and stored with {fst}.

```{r}
# document read zips and save out as fst
```

## Data Preparation

```{r}
# ds <- fst::read_fst(path = paste0(pth, "OSN_2019_fst")) %>% tibble() 
ds <- fst::read_fst(path = paste0(pth, "OSN_2020_fst")) %>% tibble() 

if(!file.exists("./data/meta.csv")){ 
  meta <- list()   # initialise meta data file
  meta$OSN_2019 <- meta_dataset_description(ds, 2019)
}else{
    meta <- readr::read_csv("./data/meta.csv")
    next_meta <- meta_dataset_description(ds, 2020)
    meta <- meta %>% left_join(next_meta)
}
meta <- meta %>% purrr::map_dfr(.f = bind_rows)
readr::write_csv(x = meta, file="./data/meta.csv")
```

Geographic references for mapping of lat/lon positions to country and/or EEZ.
Marineregions.org provides shapefile for EEZ and land boundaries (@EEZland_2020).
Land country borders are taken from the {rnaturalearth} package.

```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")
world <- world %>% select(iso_a3)

EEZ_land_shp <- list.files("../__DATA/xWorld_EEZ_and_land_v3", pattern = ".shp$", full.names=TRUE)
eez_land <- sf::read_sf(EEZ_land_shp)
eez_land <- eez_land %>% select(ISO_TER1)


EEZ_12NM_shp <- list.files("../__DATA/xWorld_12NM_v3", pattern = ".shp$", full.names=TRUE)
eez_12NM <- sf::read_sf(EEZ_12NM_shp)
eez_12NM <- eez_land %>% select(ISO_TER1)
```

```{r}
# assign unique key
ds <- ds %>% mutate(id = row_number())

# extract missing ADEPs and convert to sf pts
rq <- ds %>% select(id, icao24, origin, destination,latitude_1, longitude_1) %>%
  filter(is.na(origin)) %>% pts_latlon_to_sf(latitude_1, longitude_1, FALSE)
# rq 2020: 5,572,449

# perform spational join with 12NM territory
# runs for about 9 minutes
rq <- rq %>% 
  sf::st_join(eez_12NM)
rq <- rq %>% rename(POS1_TW = ISO_TER1)
# 2020 nas 88,215 
# remove doubles ---------------------- check todo
# another costly operation! about 10 mins
rq <- rq %>% group_by(id) %>% slice(1) %>% ungroup()

# TODO - CHECK FOR DOUBLE HITS
# rq %>% group_by(id) %>% summarise(N = n()) %>% filter(N > 1)
# resulted in 255 double hits!

# 200NM EEZ did not deliver more results --------------------------------------
# rq2 <- rq %>% filter(is.na(POS1_TW))
# rq2 <- rq2 %>% sf::st_join(eez_land)
# does not deliver any additional mapping, column left blank due to size! -----
```
rq <- rq %>% sf::st_drop_geometry()
rq %>% sf::st_set_geometry(NULL) %>% readr::write_csv("ds_origin_clean.csv")


```{r}
# load cleaner version of data
# rq <- readr::read_csv("ds_origin_clean.csv")

ds <- ds %>% left_join(rq %>% select(id, POS1_TW), by = "id")
```

tackle arrival side

```{r}
rq <- ds %>% select(id, icao24, origin, destination,latitude_2, longitude_2) %>%
  filter(is.na(destination))
# 2019 nas: 7214146
# 2020 nas: 3,973,629
rq_nope <- rq %>% filter(is.na(latitude_2) | is.na(longitude_2))
# 2019 nopes 131
# 2020 nopes 78

rq <- rq %>% filter(! id %in% rq_nope$id) %>%
  pts_latlon_to_sf(latitude_2, longitude_2, FALSE)

# perform spational join with 12NM territory
# runs for about 9 minutes
# 2019 NAs 12NM 109,152
# 2020          74,594
# no additional hits for EEZ
rq <- rq %>% 
  sf::st_join(eez_12NM)
# rename variable
rq <- rq %>% rename(POS2_TW = ISO_TER1)
## check for EEZ hits
## rq <- rq %>% sf::st_join(eez_land)
## rq %>% filter(is.na(POS2_TW) & !is.na(ISO_TER1))


# when done drop geometry
rq <- rq %>% sf::st_drop_geometry()
```

ds <- ds %>% left_join(rq %>% select(id, POS2_TW), by = "id")
sf drop geometry
ds <- ds %>% mutate(day = as.character(day)) 
ds <- ds %>% mutate(lastseen = as.character(lastseen)) 
ds <- ds %>% mutate(firstseen = as.character(firstseen)) 

fst::write_fst(ds, path = paste0(pth, "OSN_2020_clean"))

## Appendix

### Shapefile borders

Potential sources for water-land-zones boundaries

* OpenStreetMap: https://osm-boundaries.com/Map; land & water 12NM
* marineregions.org
  + EEZ in various levels, e.g. 200NM
  + EEZ plus land

### Opensky Network Data Dictionary

Description of the dataset

One file per month is provided as a csv file with the following features:

* callsign: the identifier of the flight displayed on ATC screens (usually the first three letters are reserved for an airline: AFR for Air France, DLH for Lufthansa, etc.)
* number: the commercial number of the flight, when available (the matching with the callsign comes from public open API)
* icao24: the transponder unique identification number;
* registration: the aircraft tail number (when available);
* typecode: the aircraft model type (when available);
* origin: a four letter code for the origin airport of the flight (when available);
* destination: a four letter code for the destination airport of the flight (when available);
* firstseen: the UTC timestamp of the first message received by the OpenSky Network;
* lastseen: the UTC timestamp of the last message received by the OpenSky Network;
* day: the UTC day of the last message received by the OpenSky Network;
* latitude_1, longitude_1, altitude_1: the first detected position of the aircraft;
* latitude_2, longitude_2, altitude_2: the last detected position of the aircraft.


## References
